---
title: "SIPPDATAPREP"
output: html_notebook
---

```{r}
library(tidyverse)
library(data.table)
```

Read in 2019 and 2022 datasets with selected columns

```{r}
SELECTED_COLS = c(
  # Survey Identification Variables
  "SSUID", "PNUM", "MONTHCODE", "ERELRPE", "THHLDSTATUS", 
  # Demographic + Socioeconomic Variables
  "EREFPAR", "ESEX", "TAGE", "TAGE_EHC", "ERACE", 
  "EORIGIN", "EEDUC", "TLIVQTR", "ETENURE", "ESPEAK", "RAFEVER", "EMS", "EBORNUS", "ECITIZEN", "TYRENTRY",
  "TIMSTAT", "TLANG1", "EHOWWELL", "TCEB", "TBORNPLACE", "ECERT", "EXMAR", "EVERDIV", "TRESDUR", "EED_SCRNR",
  "EEDCRED", "RGED", "EEVERET", "TAGE_FB", "EAF_HHLD", "RLNGISOL", "RFAMKIND", "RFPERSONS", "RFRELU18",
  "TEHC_REGION", "TEHC_ST", "TEHC_METRO", "EOTHSUPRT1YN", "EOTHSUPRT2YN", "EOTHSUPRT3YN", "EAWBSAFE",
  "EPNSPOUSE", "EPNPAR1", "EPNPAR2", "EJB1_JBORSE", "EJB1_CLWRK", "EJB1_UNION", "TJB1_EMPALL", "TJB1_PVOTHRC",
  # Income and Debt Variables
  "RHLTHMTH", "TPEARN", "RSSI_MNYN", "EPRLOAN1RATE", "TRENTMORT", "TINC_RENT", "TDEBT_CC", "TDEBT_ED",
  "TNETWORTH", "TVAL_RMU", "TPTOTINC", 
   # Asset Ownership and Worth Variables 
  "TVAL_STMF", "TVAL_BUS", "TVAL_RENT", "TVAL_BANK", "TVAL_BOND","TVAL_HOME", "EOWN_IRAKEO", "EOWN_THR401",
  "EOWN_PENSION", "EOWN_MM", "EOWN_CD", "TVEH_NUM", "TIRAKEOVAL",
  "TTHR401VAL", 
	#The base weight
	'WPFINWGT'
)

colTypes = c('numeric','numeric','numeric','character','character','character', 'character','numeric',
                     'numeric','character','character','character','character','character','character',
                     'character', 'character','character','character','character','character','character',
                     'character','character','character','character','character','character','character',
                     'character','character','character','character','character','character','character',
                     'character','character','character','character','character','character','character',
                     'character','character','character','character','character','character','character',
                     'character','character','character','character','character','numeric','character',
                     'numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric',
                     'numeric','numeric','numeric','numeric','numeric','numeric','character','character',
                     'character','character','character','numeric','numeric','numeric','numeric')

colClasses_named <- setNames(colTypes, SELECTED_COLS)

data2019 <- fread("/Users/vlad/Downloads/pu2019.csv", sep = ",", select = SELECTED_COLS, colClasses = colClasses_named)
data2022 <- fread("/Users/vlad/Downloads/pu2022.csv", sep = "|", select = SELECTED_COLS, colClasses = colClasses_named)

```
Combine datasets
```{r}
combined_data <- rbindlist(list(data2019, data2022))
# make OWN_STMF variable
```
Cleaning Data
  First we remove all people below 18 since they are not legally allowed to own stock.
  We then remove all retired people as they are not of interest in this survey 
  We then make Hispanic its own race by assigning it race code "5"
  We then take averages on various income variables before grouping
  each individual by their SSUID and PNUM to get one entry per person,
  as apposed to entries for each month the interview was conducted per
  person. 
  Lastly we make stock ownership a binary variable as well as make any
  empty values NA
```{r}
combined_data %>% 
  filter(TAGE >= 18 & EEVERET != "1" & !is.na(TVAL_STMF)) %>% 
  mutate(ERACE = ifelse(EORIGIN == "1", "5", ERACE)) %>%
  mutate(MEAN_TPEARN = mean(TPEARN,na.rm = TRUE)) %>%
  mutate(MEAN_TIRAKEOVAL = mean(TIRAKEOVAL,na.rm = TRUE)) %>%
  mutate(MEAN_TTHR401VAL = mean(TTHR401VAL,na.rm = TRUE)) %>%
  mutate(MEAN_TPTOTINC = mean(TPTOTINC,na.rm = TRUE)) %>%
  mutate(MEAN_TINC_RENT = mean(TINC_RENT,na.rm = TRUE)) %>%
  group_by(SSUID, PNUM) %>% 
  slice_tail(n=1) %>% 
  mutate(OWN_STMF = if_else(TVAL_STMF > 0, 1, 0)) %>% 
  mutate(across(everything(), ~ ifelse(. == "", NA, .))) 
```

Break up Dataset
```{r}
hispanic_data <- combined_data %>% 
  filter(EORIGIN == "1")
non_hispanic_data <- combined_data %>% 
  filter(EORIGIN == "2")
```


Write data out

```{r}
csv_file_path <- "/Users/vlad/Downloads/GitHub/SIPPDataAnalysis/CLEANEDSIPPDATA.csv"
write.csv(combined_data, file = csv_file_path, row.names = FALSE)
```


